---
title: "Value Chain Prioritization"
subtitle: "Expert workshop on Fonio (*Digitaria exilis*) and Jute Mallow (*Corchorus olitorius*) value chains in Northern Benin, November 2025"
author: Cory Whitney
output: html_document
bibliography: bib/refs.bib
---

<!-- add toc -->
<!-- add yaml subtitle Akkra region in Nattitengu etc.)  -->
<!-- dates of workshop -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(plotly)
library(ggrepel)
library(grid)
library(gridExtra)
```

As part of an expert workshop in November 2025 we determined potential interventions and objectives for Fonio (*Digitaria exilis*) and Jute Mallow (*Corchorus olitorius*) value chains. We conducted card-based voting (1=low, 2=medium, 3=high impact). We normalized the responses using Percentage of Maximum Possible (POMP) scoring [@cohen1999pomp] to account for participation rates and enable cross-intervention comparison. We chose the approach to simplify the rating of relationships between objectives, aligning with established participatory methods in rural development [@chambers1994participatory] and stakeholder engagement metrics [@reed2009stakeholder]. The work was facilitated by a decision modeling team, including 13 researchers *(3 academics, 3 nutritionists, 7 research institution members)*.  

The expert elicited information clarifies what actors in these value chains care about, and what they agree are possible leverage points or interventions to improve the value chains. The voting data clarifies their perception of the strength of the interaction between the interventions and their impact on the outcomes. The bubble graphs show relationships between the mean value for *impact* and *feasibility*. Income generation, for example, is considered impact whereas conflict reduction is related to feasibility. The reasoning is related to Keeney's methods for group decision making [@keeney_foundations_2013] and defining objectives [@keeney_give_2021]. This categorization separates what interventions aim to achieve (impact) from how readily they can be implemented (feasibility), supporting balanced decision-making that considers both potential benefits and practical constraints. 

```{r child="Index_fonio.Rmd"}
```

```{r child="Index_crincrin.Rmd"}
```
