---
title: "Value Chain Prioritization"
output: html_document
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(plotly)
library(ggrepel)
```

Experts identified the interventions and the objectives and then voted on the impact of each. They voted with cards representing three options, 1=low, 2=middle, and 3=high impact. We converted all responses to Percentage of Maximum Possible (POMP) scoring as applied in Participatory Rural Appraisal (PRA), community-based monitoring systems, stakeholder engagement metrics and consensus measurement in deliberative democracy. 

```{r data}
# Load the transposed matrix
# unified_matrix <- read_csv("data/interventions_matrix.csv")
unified_matrix_fonio_all <- read_csv("data/fonio_data/interventions_matrix_all.csv")
total_participants <- 55
high_vote_score <- 3
max_possible_per_intervention <- total_participants * high_vote_score 
normalized_matrix_fonio_all <- unified_matrix_fonio_all %>%
  # For score rows only, normalize across intervention columns
  mutate(across(5:13, # the rows that are vote scores
                ~ ifelse(type == "score", 
                        (.x / max_possible_per_intervention) * 10,  # Convert to 1-10 scale
                        .x)))  # Keep weight rows as-is

unified_matrix_fonio_men_boukoumbe <- read_csv("data/fonio_data/interventions_matrix_men_boukoumbe.csv")

unified_matrix_fonio_men_natitingou <- read_csv("data/fonio_data/interventions_matrix_men_natitingou.csv")

unified_matrix_fonio_women_boukoumbe <- read_csv("data/fonio_data/interventions_matrix_women_boukoumbe.csv")

unified_matrix_fonio_women_natitingou <- read_csv("data/fonio_data/interventions_matrix_women_natitingou.csv")

unified_matrix = normalized_matrix_fonio_all

# Scores were done with 1-3 votes - people who did not vote could have indicated that they are a '0' no confidence, or a 'na' that they either did not understand or did not agree. So that the total score could potentially range from n to n*3 participants. 
```

Step 1: Impact Assessment Matrix
Rate expected impact of each intervention (1-10 scale)
*10 = Very high impact, 5 = Moderate impact, 1 = No impact*

```{r unified_matrix, echo=FALSE}
source("R/create_full_heatmap_data.R")
```


Comprehensive Impact Matrix Heatmap

This shows the complete assessment - both the stakeholder priorities (weights) and intervention impacts.


```{r plot_heatmap_display_data, echo=FALSE}
source("R/plot_heatmap_display_data.R")
```

# Bubble Chart - "Impact vs Feasibility (Weighted Composite)"

This chart shows the relationship between impact and feasibility for each intervention option, with a weighted composite average score for each. Bubble size represents the total weighted composite score.

```{r bubble_graph, echo=FALSE}

# Clean intervention names for display
results <- results %>%
  mutate(intervention_clean = str_replace_all(intervention, "_", " "))

# Use ggrepel

xmin <- min(results$feasibility)
xmax <- max(results$feasibility)
ymin <- min(results$impact)
ymax <- max(results$impact)

ggplot(results, aes(x = feasibility, y = impact, size = total_score)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_text_repel(aes(label = intervention_clean), size = 3, max.overlaps = 20) +
  # # ADD strategic quadrant labels
  # annotate("text", x = 18, y = 2, label = "Easy Wins", color = "red", fontface = "bold") +
  # annotate("text", x = 18, y = 18, label = "Strategic Priorities", color = "darkgreen", fontface = "bold") +
  # annotate("text", x = 2, y = 18, label = "High Effort", color = "orange", fontface = "bold") +
  # annotate("text", x = 2, y = 2, label = "Avoid", color = "grey", fontface = "bold") +
  scale_size_continuous(range = c(3, 10), name = "Total Score") +
  labs(title = "Strategic Investment Map: Impact vs Feasibility",
       subtitle = "Coordinates show STAKEHOLDER-WEIGHTED aggregates | Size = overall value",
       x = "Feasibility (Weighted Composite)", y = "Impact (Weighted Composite)") + 
  theme_minimal() +
  xlim(xmin, xmax) + ylim(ymin, ymax)
```

# Annex - what drives the scores

```{r scores, echo=FALSE}
weight_breakdown <- score_matrix %>%
  pivot_longer(cols = -objective, names_to = "intervention", values_to = "impact_score") %>%
  left_join(objective_weights, by = "objective") %>%
  mutate(weighted_impact = impact_score * weight/100)

ggplot(weight_breakdown, aes(x = intervention, y = weighted_impact, fill = objective)) +
  geom_col() +
  labs(title = "What Drives Each Intervention's Score?",
       subtitle = "Breakdown of weighted contributions by objective") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

